An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2021-07-23 latex_files/ch2/2_1_LFI_intro.tex
create tikz figures and plots
applied fixes

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one (or many) parameter configuration(s) $\thetab^*$
that \textit{describe} the data in the best manner i.e.\ generate
samples $M_r(\thetab^*)$ that are as close as possible to the observed
data $\data$. In our case, following the Bayesian approach, we treat
the parameters of interest $\thetab$ as random variables and we try to
\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.

>>>>

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one (or many) parameter configuration(s) $\thetab^*$
that \textit{describe} the data best i.e.\ generate samples
$M_r(\thetab^*)$ that are as close as possible to the observed data
$\data$. Following the Bayesian approach, we treat the
parameters of interest $\thetab$ as random variables and we try to
\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.

### 2021-07-23 latex_files/ch2/2_2_LFI_background.tex
applied fixes

In simulator-based models we cannot evaluate the posterior
$p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
following equation allows incorporating the simulator in the place of
the likelihood,

@@

Intuitively, equation~\ref{eq:likelihood} describes that the
likelihood of a specific parameter configuration $\thetab$ is equal to
the probability that the simulator will produce outputs infinitively
close to the observations $\data$, using this configuration.

For continuous cases, the probability of replicating the data
infinitely close becomes zero. In this scenario a relaxation must be
introduced by accepting simulated data that fall close to the data
i.e. in a region $B_{d,\epsilon}(\data)$ around $\data$, where
$\epsilon > 0$. The region can be
defined as $\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$
where $d(\cdot, \cdot)$ can represent any valid distance.

Such relaxation, introduces the approximate likelihood,

@@

and consequently the approximate posterior,

>>>>

In simulator-based models it is infeasible to evaluate the posterior
$p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
following equation allows incorporating the simulator in the place of
the likelihood,

@@

Intuitively, equation~\ref{eq:likelihood} describes that the
likelihood of a specific parameter configuration $\thetab$ is equal to
the probability that the simulator will produce outputs infinitively
close to the observations $\data$, using this configuration. For
continuous cases, the probability of perfectly replicating the
observed data becomes zero. In this scenario a relaxation must be
introduced by accepting simulated data that fall close to the data
i.e. in a region $B_{d,\epsilon}(\data)$ around $\data$, where
$\epsilon > 0$. The region can be defined as
$\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$ where
$d(\cdot, \cdot)$ can represent any valid distance.

This relaxation introduces the approximate likelihood,

@@

and the approximate posterior,

### 2021-07-23 latex_files/ch2/2_3_OMC.tex
applied fixes

Before we define the likelihood approximation as introduced in the
OMC, approach lets define the indicator function based on
$\region(\yb)$. The indicator function $\indicator{\region(\yb)}(\xb)$
returns 1 if $\xb \in \region(\yb)$ and 0 otherwise. If
$d(\cdot,\cdot)$ is a formal distance, due to symmetry
$\indicator{\region(\yb)}(\xb) = \indicator{\region(\xb)}(\yb)$, so
the expressions can be used interchangeably.

@@

\noindent
Based on equation~\eqref{eq:approx_posterior} and the indicator
function as defined above~\eqref{eq:indicator}, we can approximate the
likelihood as:

\begin{gather} \label{eq:approx_likelihood_omc}
  L_{d, \epsilon}(\thetab) =
  \int_{\yb \in B_\epsilon(\data)}p(\yb|\thetab)d\yb =
  \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb\\
  \approx \frac{1}{N} \sum_i^N \indicator{\region(\data)}(\yb_i),\text{ where }
  \yb_i \sim M_r(\thetab) \label{eq:init_view}\\
  \approx \frac{1}{N} \sum_i^N \indicator{\region (\data)} (\yb_i)
  \text{ where } \yb_i = M_d(\thetab, \vb_i), \vb_i \sim p(\vb) \label{eq:alt_view}
\end{gather}
%
This approach is quite intuitive; approximating the likelihood of a
specific $\thetab$ requires sampling from the data generator and count
the fraction of samples that lie inside the area around the
observations. Nevertheless, by using the approximation of equation
\eqref{eq:init_view} we need to draw $N$ new samples for each distinct
evaluation of $L_{d,\epsilon}(\thetab)$; this makes this approach
quite inconvenient from a computational point-of-view. For this
reason, we choose to approximate the integral as in equation
\eqref{eq:alt_view}; the nuisance variables are sampled once
$\vb_i \sim p(\vb)$ and we count the fraction of samples that lie
inside the area using the deterministic simulators
$M_d(\thetab, \vb_i) \: \forall i$. Hence, the evaluation
$L_{d,\epsilon}(\thetab)$ for each different $\thetab$ does not imply
drawing new samples all over again. Based on this approach, the
unnormalised approximate posterior can be defined as:

@@

The posterior approximation in \eqref{eq:approx_posterior_omc} does not
provide any obvious way for drawing samples. In fact, the set
$\mathcal{S}_i = \{ \thetab: M_d(\thetab, \vb_i) \in \region(\data) \}$ can
represent any arbitrary shape in the D-dimensional Euclidean space; it
can be non-convex, can contain disjoint sets of $\thetab$ etc. We need
some further simplification of the posterior for being able to draw
samples from it.

As a side-note, weighted sampling could be performed in a
straightforward fashion with importance sampling. Using the prior as
the proposal distribution $\thetab_i \sim p(\thetab)$ and we can
compute the weight as
$w_i = \frac{L_{d,\epsilon}(\thetab_i)}{p(\thetab_i)}$, where
$L_{d,\epsilon}(\thetab_i)$ is computed with the expression
\eqref{eq:approx_likelihood_omc}. This approach has the same drawbacks as
ABC rejection sampling; when the prior is wide or the dimensionality
$D$ is high, drawing a sample with non-zero weight is rare, leading to
either poor Effective Sample Size (ESS) or huge execution time.

The OMC proposes a quite drastic simplification of the posterior; it
squeezes all regions $\mathcal{S}_i$ into a single point
$\thetab_i^* \in \mathcal{S}_i$ attaching a weight $w_i$ proportional
to the volume of $\mathcal{S}_i$. For obtaining a
$\thetab_i^* \in \mathcal{S}_i$, a gradient based optimiser is used
for minimising $g_i(\thetab) = d(\data, f_i(\thetab))$ and the
estimation of the volume of $\mathcal{S}_i$ is done using the Hessian
approximation $\hess_i \approx \jac_i^{*T}\jac_i^*$, where $\jac_i^*$
is the Jacobian matrix of $g_i(\thetab)$ at $\thetab_i^*$. Hence,

>>>>

For describing the likelihood approximation introduced in OMC, it is
convenient to firstly define the indicator function based on
$\region(\yb)$. The indicator function $\indicator{\region(\yb)}(\xb)$
returns 1 if $\xb \in \region(\yb)$ and 0 otherwise. If
$d(\cdot,\cdot)$ is a formal distance, due to symmetry
$\indicator{\region(\yb)}(\xb) = \indicator{\region(\xb)}(\yb)$, so
the expressions can be used interchangeably.

@@

\noindent
Based on equation~\eqref{eq:approx_likelihood} and the indicator
function~\eqref{eq:indicator}, we can approximate the likelihood as:

\begin{gather} \label{eq:approx_likelihood_omc}
  L_{d, \epsilon}(\thetab) =
  \int_{\yb \in B_\epsilon(\data)}p(\yb|\thetab)d\yb =
  \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb\\
  \approx \frac{1}{N} \sum_i^N \indicator{\region(\data)}(\yb_i),\text{ where }
  \yb_i \sim M_r(\thetab) \label{eq:init_view}\\
  \approx \frac{1}{N} \sum_i^N \indicator{\region (\data)} (\yb_i)
  \text{ where } \yb_i = M_d(\thetab, \vb_i), \vb_i \sim p(\vb) \label{eq:alt_view}
\end{gather}
%
The procedure is quite intuitive; approximating the likelihood of a
specific $\thetab$ requires sampling from the data generator and count
the fraction of samples that lie inside the area around the
observations. Nevertheless, by using the approximation of equation
\eqref{eq:init_view} we need to draw $N$ new samples for each distinct
evaluation of $L_{d,\epsilon}(\thetab)$; this makes this approach
quite inconvenient from a computational point-of-view. For this
reason, OMC approximates the integral as in equation
\eqref{eq:alt_view}; the nuisance variables, $\vb_i \sim p(\vb)$, are
sampled once and we count the fraction of samples that lie inside the
area using the deterministic simulators
$M_d(\thetab, \vb_i) \: \forall i$. Hence, the evaluation
$L_{d,\epsilon}(\thetab)$ for each different $\thetab$ does not imply
drawing new samples all over again. Based on this approach, the
unnormalised approximate posterior can be defined as:

@@

The posterior approximation in \eqref{eq:approx_posterior_omc} does
not provide any convenient way for drawing samples.\footnote{Weighted
  sampling could be performed in a straightforward fashion with
  importance sampling, using the prior as the proposal distribution
  $\thetab_i \sim p(\thetab)$. However, this approach has the same
  drawbacks as ABC rejection sampling; when the prior is wide or the
  dimensionality $D$ is high, drawing a sample with non-zero weight is
  rare, leading to either poor Effective Sample Size (ESS) or huge
  execution time.} In fact, the set
$\mathcal{S}_i = \{ \thetab: M_d(\thetab, \vb_i) \in \region(\data)
\}$ can represent any arbitrary shape in the D-dimensional Euclidean
space; it can be non-convex, can contain disjoint sets of $\thetab$
etc. We need some further simplification of the posterior for being
able to draw samples from it.

OMC proposes a quite drastic simplification of the posterior; it
squeezes all regions $\mathcal{S}_i$ into a single point
$\thetab_i^* \in \mathcal{S}_i$ attaching a weight $w_i$ proportional
to the volume of $\mathcal{S}_i$. For obtaining a
$\thetab_i^* \in \mathcal{S}_i$, a gradient based optimiser is used
for minimising $g_i(\thetab) = d(\data, f_i(\thetab))$ and the
estimation of the volume of $\mathcal{S}_i$ is done using the Hessian
approximation $\hess_i \approx \jac_i^{*T}\jac_i^*$, where $\jac_i^*$
is the Jacobian matrix of $g_i(\thetab)$ at $\thetab_i^*$. Hence,

### 2021-07-23 latex_files/ch2/2_5_ROMC_algorithmic.tex
create tikz figures and plots
applied fixes

Here we present ROMC as a sequence of algorithmic steps. The reader
may think of the algorithmic description as the backbone of the
implementation that will follow. At a high-level, ROMC can be split
into the training and the inference part.

\subsubsection*{Training part}
\noindent
At the training (fitting) part, the goal is the estimation of the
proposal regions $q_i$. The tasks are (a) sampling the nuisance
variables $\vb_i \sim p(\vb)$ (b) defining the optimisation problems
$\min_{\thetab} \: g_i(\thetab)$ (c) obtaining $\thetab_i^*$ (d)
checking whether $d_i^* \leq \epsilon$ and (e) building the bounding
box for obtaining the proposal region $q_i$. If gradients are
available, using a gradient-based method is advised for obtaining
$\thetab_i^*$ much faster. Providing $\nabla_{\thetab} g_i$ in
closed-form provides an upgrade in both accuracy and efficiency; If
closed-form description is not available, approximate gradients with
finite-differences requires two evaluations of $g_i$ for \emph{every}
parameter $\theta_d$, which can only work in low-dimensional
problems. When gradients are not available, the Bayesian optimisation
paradigm exists as an alternative solution. In this scenario, the
training part becomes slower due to fitting of the surrogate model and
the blind optimisation steps. Nevertheless, the subsequent task of
computing the proposal region $q_i$ becomes faster since $\hat{d}_i$
can be used instead of $g_i$. The algorithms~\ref{alg:training_GB}
and~\ref{alg:training_GP} present the above procedure.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$ and (c) evaluating the
unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Computing an
expectation can be done easily after weighted samples are obtained
using the equation~\ref{eq:expectation}, so we do not discuss it
separately.

\noindent
Sampling is performed by getting $n_2$ samples from each proposal
distribution $q_i$. For each sample $\thetab_{ij}$, the indicator
function is evaluated $\indicator{\regioni(\data)}(\thetab_{ij})$ for
checking if it lies inside the acceptance region. If so the
corresponding weight is computed as in \eqref{eq:sampling}. As before,
if a surrogate model $\hat{d}$ is available, it can be utilised for
evaluating the indicator function. At the sampling task, the
computational benefit of using the surrogate model is more valuable
compared to the evaluation of the posterior, because the indicator
function must be evaluated for a total of $n_1 \times n_2$ points.

\noindent
Evaluating the unnormalised posterior requires only the
deterministic functions $g_i$ and the prior distribution $p(\thetab)$;
there is no need for solving the optimisation problems and building
the proposal regions. The evaluation requires iterating over all $g_i$
and evaluating the distance from the observed data. In contrast, using
the GP approach, the optimisation part should be performed first for
fitting the surrogate models $\hat{d}_i(\thetab)$ and evaluate the
indicator function on them. This provides an important speed-up,
especially when running the simulator is computationally
expensive.

\noindent
In summary, we can state that the choice of using a Bayesian
optimisation approach provides a significant speed-up in the inference
part with the cost of making the training part slower and a possible
approximation error. It is typical in many Machine-Learning use cases,
being able to provide enough time and computational resources for the
training phase, but asking for efficiency in the inference
part.

@@

% \begin{minipage}{0.46\textwidth}
% \begin{algorithm}[H]
%     \centering
%     \caption{Evaluate unnormalised posterior - Gradient approach. Requires $g_i(\theta), p(\theta)$}\label{alg:posterior_GB}
%     \begin{algorithmic}[1]
%       \State $k \leftarrow 0$
%         \For {$i \gets 1 \textrm{ to } n_1$}
%           \If {$g_i(\theta) > \epsilon$}
%             \State $k \leftarrow k + 1$
%           \EndIf
%           \EndFor
%       \Return{$kp(\theta)$}
%     \end{algorithmic}
% \end{algorithm}
% \end{minipage}
% \hfill
% \begin{minipage}{0.46\textwidth}
% \begin{algorithm}[H]
%     \centering
%     \caption{Evaluate unnormalised posterior - GP approach. Requires $\hat{d}_i(\theta), p(\theta)$}\label{alg:posterior_GP}
%     \begin{algorithmic}[1]
%       \State $k \leftarrow 0$
%         \For {$i \gets 1 \textrm{ to } n_1$}
%           \If {$d_i(\theta) > \epsilon$}
%             \State $k \leftarrow k + 1$
%           \EndIf
%           \EndFor
%       \Return{$kp(\theta)$}
%     \end{algorithmic}
% \end{algorithm}
% \end{minipage}

% \begin{minipage}{0.46\textwidth}
\begin{algorithm}[H]
    \centering
    \caption{Sampling. Requires a function of distance $(g_i(\theta)$ or $\hat{d}_i$ or $\hat{g}_i), p(\theta), q_i$}\label{alg:sampling_GB}
    \begin{algorithmic}[1]
      \For {$i \gets 1 \textrm{ to } n_1$}
      \For {$j \gets 1 \textrm{ to } n_2$}
          \State $\thetab_{ij} \sim q_i$
          \If {$g_i(\thetab_{ij}) > \epsilon$}
            \State Reject $\theta_{ij}$
          \Else {}
            \State $w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}$
            \State Accept $\thetab_{ij}$, with weight $w_{ij}$
          \EndIf
      \EndFor
      \EndFor
    \end{algorithmic}
\end{algorithm}
% \end{minipage}
% \hfill
% \begin{minipage}{0.46\textwidth}
% \begin{algorithm}[H]
%     \centering
%     \caption{Sampling - GP approach. Requires $\hat{d}_i(\theta), p(\theta), q_i$}\label{alg:sampling_GP}
%     \begin{algorithmic}[1]
%       \For {$i \gets 1 \textrm{ to } n_1$}
%       \For {$j \gets 1 \textrm{ to } n_2$}
%           \State $\theta_{ij} \sim q_i$
%           \If {$\hat{d}_i(\theta_{ij}) > \epsilon$}
%             \State Reject $\theta_{ij}$
%           \Else {}
%             \State $w_{ij} = \frac{p(\theta_{ij})}{q(\theta_{ij})}$
%             \State Accept $\theta_{ij}$, with weight $w_{ij}$
%           \EndIf
%       \EndFor
%       \EndFor
%     \end{algorithmic}
% \end{algorithm}
% \end{minipage}

>>>>

For designing an extendable implementation of ROMC, we firstly define
ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
serves as the driver for the implementation at \pkg{ELFI} that
follows. At a high level, ROMC can be split into the training and the
inference part. In general terms, the training part covers all the
appropriate steps for estimating the proposal regions and the
inference part for getting weighted
samples. Algorithm~\ref{alg:romc_algorithm} defines formally ROMC;
Steps 2-10 (before the vertical line) form the training part, while
steps 12 and 13 the inference part.

\begin{algorithm}[!ht]
	\caption{ROMC}\label{alg:romc_algorithm}
	\begin{algorithmic}[1]
    \Procedure{ROMC}{$M_r(\thetab)$}
    \For{$i \gets 1 \textrm{ to } n_1$}
      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define deterministic distance function}
      \State Obtain $\thetab_i^*, d_i^*$ \Comment{Solve optimization problem}
    \EndFor
    \State $\text{L} \gets \{d_{i, i = 1, ..., n_1}: d_i^* < \epsilon \} $ \Comment{List with filtered solutions}
    \For{$i \gets 1 \textrm{ to } len(L)$}
      \State Construct $q_i$ \Comment{Compute proposal area}
      \State Compute $\Tilde{d}_i$ \Comment{Fit surrogate model in $q_i$}
      \EndFor
    \\\hrulefill
    \State Draw $\{w_{ij}, \thetab_{ij}\}$ \Comment{Draw $n_2$ samples per $q_i$}
    \State Draw $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}    
    \EndProcedure
	\end{algorithmic}
\end{algorithm}

\subsubsection*{Training part}
\noindent
At the training (fitting) part, the goal is the estimation of the
proposal regions $q_i$. The tasks are; (a) sampling the nuisance
variables $\vb_i \sim p(\vb)$ for obtaining the deterministic
functions $g_i(\thetab)$, (b) defining the optimisation problems
$\min_{\thetab} \: g_i(\thetab)$, (c) obtaining $\thetab_i^*$ and
$d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and (e) building
the bounding box for obtaining the proposal region $q_i$. If gradients
are available, using a gradient-based method is advised for obtaining
$\thetab_i^*$ much faster. In case $\nabla_{\thetab} g_i$ is available
in closed-form, the procedure is upgraded in terms of both accuracy
and efficiency; If closed-form description is not available, gradients
approximation with finite-differences requires two evaluations of
$g_i$ for \emph{each} parameter $\theta_d$, which can only work in
low-dimensional problems. If $g_i(\thetab)$ is not differentiable, the
Bayesian Optimisation can be used as an alternative solution. In this
scenario, the training part becomes slower due to fitting of the
surrogate model and the blind optimisation
steps. Algorithm~\ref{alg:training_GB} defines formally the steps of a
gradient-based optimiser and algorithm~\ref{alg:training_GP} the steps
of the Bayesian Optimisation. After obtaning the optimal points
$\thetab^*$, the training part gets completed by estimating the
proposal regions. Algorithm~\ref{alg:region_construction} describes
the line search approach for finding the region's boundaries.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$ or (c) evaluating the
unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Computing an
expectation can be done easily after weighted samples are obtained
using the equation~\ref{eq:expectation}, so we do not discuss it
separately.

\noindent
Sampling is performed by getting $n_2$ samples from each proposal
distribution $q_i$. For each sample $\thetab_{ij}$, the indicator
function is evaluated $\indicator{\regioni(\data)}(\thetab_{ij})$ for
checking if it lies inside the acceptance region. If this is the case,
the corresponding weight is computed as in \eqref{eq:sampling}. As
before, if a surrogate model $\hat{d}$ is available, it can be
utilised for evaluating the indicator function.

\noindent
Evaluating the unnormalised posterior requires the deterministic
functions $g_i$ and the prior distribution $p(\thetab)$\footnote{There
  is no need for solving the optimisation problems and building the
  proposal regions.}. The evaluation requires iterating over all $g_i$
and evaluating the distance from the observed data.

@@

\begin{algorithm}[H]
    \centering
    \caption{Sampling. Requires a function of distance $(g_i(\theta)$ or $\hat{d}_i$ or $\hat{g}_i), p(\theta), q_i$}\label{alg:sampling_GB}
    \begin{algorithmic}[1]
      \For {$i \gets 1 \textrm{ to } n_1$}
      \For {$j \gets 1 \textrm{ to } n_2$}
          \State $\thetab_{ij} \sim q_i$
          \If {$g_i(\thetab_{ij}) > \epsilon$}
            \State Reject $\theta_{ij}$
          \Else {}
            \State $w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}$
            \State Accept $\thetab_{ij}$, with weight $w_{ij}$
          \EndIf
      \EndFor
      \EndFor
    \end{algorithmic}
\end{algorithm}

### 2021-07-23 latex_files/ch3/3_2_1_general_design.tex
applied fixes

The overview of our implementation is illustrated in
figure~\ref{fig:romc_overview}; one may interpret
figure~\ref{fig:romc_overview} as a depiction of the main class and
the ellipses are the main routines of the class. Following
\proglang{Python}'s naming principles, the methods starting with an
underscore (green ellipses) represent internal (private) functions,
whereas the rest (blue ellipses) are the public methods that the user
interacts with.

Figure~\ref{fig:romc_overview} splits ROMC into the training, the
inference and the evaluation part. The training part includes all the
steps until the computation of the proposal regions i.e. sampling the
nuisance variables, defining the optimization problems, solving them,
constructing the regions and fitting local surrogate models. The
inference part comprises of evaluating the unnormalized posterior (and
the normalized when is possible), sampling and computing an
expectation. From an algorithmic point of view, the training and the
inference parts are \emph{isolated}. We also provide some utilities
for inspecting the training process, such as plotting the histogram of
the final distances or visualizing the constructed bounding
boxes. Finally, in the evaluation part, we provide two methods for
evaluating the inference; (a) computing the Effective Sample Size
(ESS) of the samples and (b) measuring the divergence between the
approximate posterior the ground-truth, if the latter is
available.\footnote{Normally, the ground-truth posterior is not
  available; However, this functionality is useful in cases where the
  posterior can be computed numerically or with an alternative method
  (i.e.\ ABC Rejection Sampling), and we would like to measure the
  discrepancy between the two approximations.}

% \begin{figure}[h]
%     \begin{center}
% \includegraphics[width=0.5\textwidth]{./latex_files/graphs/ROMC.png}
%     \end{center}
%     \caption[Overview of the ROMC implementation.]{Overview of the
%       ROMC implementation. The training part follows a sequential
%       pattern; the functions in the green ellipses must be called in a
%       sequential fashion for completing the training part and define
%       the posterior distribution. The functions in blue ellipses are
%       the functionalities provided to the user.}
%     \label{fig:romc_overview}
% \end{figure}

@@

% \begin{figure}[!h]
%     \begin{center}
% \includegraphics[width=0.7\textwidth]{./latex_files/images/chapter3/example_gt.png}
%     \end{center}
%   \caption[Ground-truth posterior distribution of the simple 1D
% example.]{Ground-truth posterior distribution of the simple 1D
% example.}
%   \label{fig:example_gt}
% \end{figure}

>>>>

The overview of our implementation is illustrated in
figure~\ref{graph:romc_overview}; one may interpret
figure~\ref{graph:romc_overview} as a depiction of the main class and
the ellipses are the main routines of the class. Following
\proglang{Python}'s naming principles, the methods starting with an
underscore (green ellipses) represent internal (private) functions,
whereas the rest (blue ellipses) are the public methods that the user
interacts with.

\input{latex_files/graphs/romc_graph}

In figure~\ref{graph:romc_overview}, it can be easily observed that the
implementation follows consistently the algorithmic description
described above, which follows the partition into training, inference
and evaluation. The training part includes all the steps until the
computation of the proposal regions i.e. sampling the nuisance
variables, defining the optimization problems, solving them,
constructing the regions and fitting local surrogate models. The
inference part comprises of evaluating the unnormalized posterior (and
the normalized when is possible), sampling and computing an
expectation. From an algorithmic point of view, the training and the
inference parts are \emph{isolated}. We also provide some utilities
for inspecting the training process, such as plotting the histogram of
the final distances or visualizing the constructed bounding
boxes. Finally, in the evaluation part, we provide two methods for
evaluating the inference; (a) computing the Effective Sample Size
(ESS) of the samples and (b) measuring the divergence between the
approximate posterior the ground-truth, if the latter is
available.\footnote{Normally, the ground-truth posterior is not
  available; However, this functionality is useful in cases where the
  posterior can be computed numerically or with an alternative method
  (i.e.\ ABC Rejection Sampling), and we would like to measure the
  discrepancy between the two approximations.}

@@

### 2021-07-23 paper.tex
applied fixes

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{A Parallelized and Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
\Plaintitle{A Parallelized and Extendable Python Implementation of ROMC}
\Shorttitle{A Parallelized and Extendable \proglang{Python} Implementation of ROMC}

@@

For developing an extenbable implementation of ROMC, we initially
translate ROMC's mathematic description to an algorithmic one. The
pure algorithmic view serves as the driver for designing the actual
implementation at \pkg{ELFI}.

\input{latex_files/graphs/romc_graph}

\subsection{Algorithmic view of ROMC} % better titl
\input{latex_files/ch2/2_5_ROMC_algorithmic.tex}

>>>>

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
\Plaintitle{An Extendable Python Implementation of ROMC}
\Shorttitle{An Extendable \proglang{Python} Implementation of ROMC}

@@

\input{latex_files/ch2/2_5_ROMC_algorithmic.tex}
